\begin{thebibliography}{99}

\bibitem{ghostplc}
A.Abbasi, M.Hashemi, E.Zambon, S.Etalle,
``Stealth Low-Level Manipulation of Programmable Logic Controllers I/O By Pin Control Exploitation'',
TODO Complete citation CRITIS Conference 2016.

\bibitem{pinctrl}
``Pin Control Subsystem'',
\url{https://www.kernel.org/doc/Documentation/pinctrl.txt}.

\bibitem{stuxnet}
N.Falliere, L.O Murchu, E.Chien,
``W32.Stuxnet  Dossier'',
Version 1.4, Febr. 2011.
Online: \url{http://www.symantec.com/content/en/us/enterprise/media/security_response/whitepapers/w32_stuxnet_dossier.pdf}.

\bibitem{io-command}
R.Grandgenett, W.Mahoney, R.Gandhi,
``Authentication Bypass and Remote Escalated I/O Command Attacks'',
CISR '15: Proceedings of the 10th Annual Cyber and Information Security Research Conference,
Oak Ridge, Tennesse (USA), April 7-9, 2015,
\doi{10.1145/2746266.2746268}.

\bibitem{taxonomy}
D.Papp, Z.Ma, L.Buttyan,
``Embedded systems security: Threats, vulnerabilities, and attack taxonomy''
Privacy, Security and Trust (PST) 13th Annual Conference,
Izmir (Turkey), July 21-23, 2015
pp.\ 145-152,
\doi{10.1109/PST.2015.7232966}.

\bibitem{firmware-mod}
Z.Basnight, J.Butts, J.Lopez Jr., T.Dube,
``Firmware modification attacks on programmable logic controllers'',
International Journal of Critical Infrastructure Protection,
Volume 6, Issue 2, June 2013,
pp.\ 76-84,
\doi{10.1016/j.ijcip.2013.04.004}.

\bibitem{ethernet-vuln}
D.Peck, D.Peterson,
``Leveraging ethernet card vulnerabilities in field devices'',
Proceedings of the SCADA Security Scientific Symposium,
Miami Beach, Florida (USA), Jan. 18-19, 2009,
pp.\ 1-19.

\bibitem{print-vuln}
A.Cui, M.Costello, S.J.Stolfo,
``When Firmware Modifications Attack: A Case Study of Embedded Exploitation'',
20th Annual Network \& Distributed System Security Symposium,
San Diego, California (USA), Febr. 24-27, 2013,
\doi{10.7916/D8P55NKB}.

\bibitem{dynamic-payload}
S.McLaughlin,
``On Dynamic Malware Payloads Aimed at Programmable Logic Controllers'',
In 6th USENIX Workshop on Hot Topics in Security,
2011.

\bibitem{sabot}
S.McLaughlin, P.McDaniel,
``SABOT: Specification-based Payload Generation for Programmable Logic Controllers'',
CCS '12: Proceedings of the 2012 ACM conference on Computer and Communications Security,
New York (USA), Oct. 16-18, 2012,
pp.\ 439-449,
\doi{10.1145/2382196.2382244}.

\bibitem{siemens-s7}
D.Beresford,
``Exploiting Siemens Simatic S7 PLCs'',
Black Hat USA 2011,
Las Vegas, Nevada (USA), Aug. 3-4, 2011.
Online: \url{https://media.blackhat.com/bh-us-11/Beresford/BH_US11_Beresford_S7_PLCs_WP.pdf}.

\bibitem{plc-network}
J.Klick, S.Lau, D.Marzin, J.Malchow, V.Roth,
``Internet-facing PLCs as a Network Backdoor'',
Proceedings of IEEE Conference on Communications and Network Security (CNS),
Florence (Italy), Sept. 28-30, 2015,
pp.\ 524-532,
\doi{10.1109/CNS.2015.7346865}.

\bibitem{abb-codesys}
ICS-CERT,
``ABB AC500 PLC Webserver CoDeSys Vulnerability'',
April 30, 2013.
Online: \url{https://ics-cert.us-cert.gov/advisories/ICSA-12-320-01}.

\bibitem{codesys-server}
ICS-CERT,
``3S CODESYS Gateway-Server Vulnerabilities (Update A)'',
March 13, 2014.
Online: \url{https://ics-cert.us-cert.gov/advisories/ICSA-13-050-01A}.

\bibitem{schneider-bof}
ICS-CERT,
``Schneider Electric Modicon M340 Buffer Overflow Vulnerability'',
Dec. 17, 2015.
Online: \url{https://ics-cert.us-cert.gov/advisories/ICSA-15-351-01}.

\bibitem{rockwell-vuln}
ICS-CERT,
``Rockwell Automation Micrologix 1100 and 1400 PLC Systems Vulnerabilities (Update A)'',
Oct. 27, 2015.
Online: \url{https://ics-cert.us-cert.gov/advisories/ICSA-15-300-03A}.

\bibitem{rockwell-vuln2}
ICS-CERT,
``Rockwell Automation MicroLogix 1100 PLC Overflow Vulnerability'',
Jan. 26, 2016.
Online: \url{https://ics-cert.us-cert.gov/advisories/ICSA-16-026-02}.

\bibitem{elcsoft-vuln}
ICS-CERT,
``Eaton ELCSoft Programming Software Memory Vulnerabilities'',
June 30, 2016.
Online: \url{https://ics-cert.us-cert.gov/advisories/ICSA-16-182-01}.

\bibitem{jop}
P.Chen, X.Xing, B.Mao, L.Xie,
``Return-oriented rootkit without returns (on the x86)''
ICICS 2010: 12th International Conference on Information and Communications Security,
Barcelona (Spain), Dec. 15-17, 2010,
pp.\ 340-354,
\doi{10.1007/978-3-642-17650-0_24}.

\bibitem{no-ret}
S.Checkoway, L.Davi, A.Dmitrienko, A.Sadeghi, H.Shacham, M.Winandy,
``Return-Oriented Programming without Returns'',
CCS '10: Proceedings of the 17th ACM conference on Computer and Communications Security,
Chicago, Illinois (USA), Oct. 4-8, 2010,
\doi{10.1145/1866307.1866370}.

\bibitem{scada-attacks}
R.E.Johnson,
``Survey of SCADA Security Challenges and Potential Attack Vectors'',
ICITST 2010: International Conference for Internet Technology and Secured Transactions,
London (UK), Nov. 8-11, 2010,
pp.\ 80-85.

\bibitem{scada-attacks2}
B.Miller, D.Rowe,
``A survey of SCADA and critical infrastructure incidents'',
RIIT '12: Proceedings of the 1st Annual conference on Research In Information Technology,
Calgary, Alberta (Canada), Oct. 10-13, 2012,
pp.\ 51-56,
\doi{10.1145/2380790.2380805}.

\bibitem{plc-security}
G.P.H.Sandaruwan, P.S.Ranaweera, V.A.Oleshchuk,
``PLC Security and Critical Infrastructure Protection'',
ICIIS 2013: IEEE 8th International Conference on Industrial and Information Systems,
Peradeniya (Sri Lanka), Dec. 17-20, 2013,
pp.\ 81-85,
\doi{10.1109/ICIInfS.2013.6731959}.

\bibitem{stuxnet-defense}
A.Clark, Q.Zhu, R.Poovendran, T.Başar,
``An Impact-Aware Defense against Stuxnet'',
ACC 2013: 1st American Control Conference,
Washington, DC (USA), June 17-19, 2013,
pp.\ 4140-4147,
\doi{10.1109/ACC.2013.6580475}.

\bibitem{semantic-defense}
D.Hadžiosmanović, R.Sommer, E.Zambon, P.H.Hartel,
``Through the Eye of the PLC: Semantic Security Monitoring for Industrial Processes'',
ACSAC '14: 30th Annual Computer Security Applications Conference,
New Orleans, LA (USA), Dec. 08-12, 2014,
pp.\ 126-135,
\doi{10.1145/2664243.2664277}.

\bibitem{confirm}
X.Wang, C.Konstantinou, M.Maniatakos, R.Karri, S.Lee, P.Robison, P.Stergiou, S.Kim,
``Malicious Firmware Detection with Hardware Performance Counters'',
IEEE Transactions on Multi-Scale Computing Systems,
Vol.\ 2, Issue 3,
July-Sept. 2016,
pp.\ 160-173,
\doi{10.1109/TMSCS.2016.2569467}.

\bibitem{trustzone}
ARM Limited,
``ARM Security Technology - Building a Secure System using TrustZone® Technology'',
2009.
Online: \url{http://infocenter.arm.com/help/topic/com.arm.doc.prd29-genc-009492c/PRD29-GENC-009492C_trustzone_security_whitepaper.pdf}.

\bibitem{trustlite}
P.Koeberl, S.Schulz, A.Sadeghi, V.Varadharajan,
``TrustLite: A Security Architecture for Tiny Embedded Devices'',
EuroSys '14: Proceedings of the Ninth European Conference on Computer Systems,
Amsterdam (Netherlands), April 13-16, 2014,
\doi{10.1145/2592798.2592824}.

\bibitem{tpm2}
A.Fuchs, C.Krau{\ss}, J.Repp,
``Advanced Remote Firmware Upgrades Using TPM 2.0'',
in the book ``ICT Systems Security and Privacy Protection: 31st IFIP TC 11 International Conference, SEC 2016, Ghent, Belgium, May 30 - June 1, 2016, Proceedings''
edited by J.Hoepman, S.Katzenbeisser,
Springer International Publishing, 2016,
pp.\ 276-289,
\doi{10.1007/978-3-319-33630-5_19}.

\bibitem{blockchain}
B.Lee, J.Lee,
``Blockchain-based secure firmware update for embedded devices in an Internet of Things environment'',
The Journal of Supercomputing, 2016,
pp.\ 1-16,
\doi{10.1007/s11227-016-1870-0}.

\bibitem{logic-analytics}
S.Zonouz, J.Rrushi, S.McLaughlin,
``Detecting Industrial Control Malware Using Automated PLC Code Analytics'',
IEEE Security \& Privacy,
Vol.\ 12, Issue 6,
Nov.-Dec. 2014,
pp.\ 40-47,
\doi{10.1109/MSP.2014.113}.

\bibitem{hypervisor-control}
L.Garcia, S.Zonouz, D.Wei, L.P.de Aguiar,
``Detecting PLC Control Corruption via On-Device Runtime Verification'',
Resilience Week (RWS),
Chicago, Illinois (USA), Aug. 16-18, 2016,
pp.\ 67-72,
\doi{10.1109/RWEEK.2016.7573309}.

\bibitem{symbiotes}
A.Cui, S.J.Stolfo,
``Defending Embedded Systems with Software Symbiotes'',
in the book ``Recent Advances in Intrusion Detection: 14th International Symposium, RAID 2011, Menlo Park, CA, USA, September 20-21, 2011. Proceedings'',
edited by R.Sommer, D.Balzarotti, G.Maier,
Springer Berlin Heidelberg, 2011,
pp.\ 358-377,
\doi{10.1007/978-3-642-23644-0_19}.

\bibitem{swatt}
A.Seshadri, A.Perrig, L.Doom, P.Khosla,
``SWATT: SoftWare-based ATTestation for embedded devices'',
Proceedings of IEEE Symposium on Security and Privacy,
Oakland, California (USA), May 9-12, 2004,
pp.\ 272-282,
\doi{10.1109/SECPRI.2004.1301329}.

\bibitem{swatt-difficulty}
C.Castelluccia, A.Francillon, D.Perito, C.Soriente,
``On the difficulty of software-based attestation of embedded devices'',
CCS'09: Proceedings of the 16th ACM Conference on Computer and Communications Security,
Chicago, Illinois (USA), Nov. 9-13, 2009,
pp.\ 400-409,
\doi{10.1145/1653662.1653711}.

\bibitem{power-fingerprinting}
C.A.Gonzalez, A.Hinton,
``Detecting Malicious Software Execution in Programmable Logic Controllers Using Power Fingerprinting'',
in the book ``Critical Infrastructure Protection VIII: 8th IFIP WG 11.10 International Conference, ICCIP 2014, Arlington, VA, USA, March 17-19, 2014, Revised Selected Papers'',
edited by J.Butts, S.Shenoi,
Springer Berlin Heidelberg, 2014,
pp.\ 15-27,
\doi{10.1007/978-3-662-45355-1_2}.

\bibitem{autoscopy}
J.Reeves, A.Ramaswamy, M.Locasto, S.Bratus, S.Smith,
``Intrusion detection for resource-constrained embedded control systems in the power grid'',
International Journal of Critical Infrastructure Protection, 2012,
Vol.\ 5, No.\ 2,
pp.\ 74-83,
\doi{10.1016/j.ijcip.2012.02.002}.

\bibitem{disarm}
J.Habibi, A.Panicker, A.Gupta, E.Bertino,
``DisARM: Mitigating Buffer Overflow Attacks on Embedded Devices'',
in the book ``Network and System Security: 9th International Conference, NSS 2015, New York, NY, USA, November 3-5, 2015, Proceedings'',
edited by M.Qiu, S.Xu, M.Yung, H.Zhang,
Springer International Publishing, 2015,
pp.\ 112-129,
\doi{10.1007/978-3-319-25645-0_8}.

\bibitem{hardware-ibmac}
A.Francillon, D.Perito, C.Castelluccia,
``Defending Embedded Systems Against Control Flow Attacks'',
SecuCode '09: Proceedings of the first ACM workshop on Secure execution of untrusted code,
Chicago, Illinois (USA), Nov. 9, 2009,
pp.\ 19-26,
\doi{10.1145/1655077.1655083}.

\bibitem{ocfmm}
F.Abad, J.Woude, Y.Lu, S.Bak, M.Caccamo, L.Sha, R.Mancuso, S.Mohan,
``On-Chip Control Flow Integrity Check for Real Time Embedded Systems'',
2013 IEEE 1st International Conference on Cyber-Physical Systems, Networks, and Applications (CPSNA),
Taipei, Taiwan, Aug. 19-20, 2013,
pp.\ 26-31,
\doi{10.1109/CPSNA.2013.6614242}.

\bibitem{fine-grained}
L.Davi, P.Koeberl, A.Sadeghi,
``Hardware-Assisted Fine-Grained Control-Flow Integrity: Towards Efficient Protection of Embedded Systems Against Software Exploitation'',
DAC '14: Proceedings of the 51st Annual Design Automation Conference,
San Francisco, California (USA), June 1-5, 2014,
pp.\ 133:1-133:6,
\doi{10.1109/DAC.2014.6881460}.

\bibitem{hafix}
L.Davi, M.Hanreich, D.Paul, A.Sadeghi, P.Koeberl, D.Sullivan, O.Arias, Y.Jin,
``HAFIX: Hardware-Assisted Flow Integrity Extension'',
DAC '15: Proceedings of the 52nd Annual Design Automation Conference,
San Francisco, California (USA), June 7-11, 2015,
pp.\ 74:1-74:6,
\doi{10.1145/2744769.2744847}.

\bibitem{bb-cfi}
S.Das, W.Zhang, Y.Liu,
``A Fine-Grained Control Flow Integrity Approach Against Runtime Memory Attacks for Embedded Systems'',
IEEE Transactions on Very Large Scale Integration (VLSI) Systems,
Vol.\ 24, No.\ 11,
Nov. 2016,
pp.\ 3193-3207,
\doi{10.1109/TVLSI.2016.2548561}.

\bibitem{cflat}
T.Abera, N.Asokan, L.Davi, J.Ekberg, T.Nyman, A.Paverd, A.Sadeghi, G.Tsudik,
``C-FLAT: Control-FLow ATtestation for Embedded Systems Software'',
CCS '16: Proceedings of the 23rd ACM Conference on Computer and Communications Security,
Vienna (Austria), Oct. 24-28, 2016,
pp.\ 743-754,
\doi{10.1145/2976749.2978358}.

\bibitem{smm-rootkit}
S.Embleton, S.Sparks, C.C.Zou,
``SMM rootkit: a new breed of OS independent malware'',
Security and Communication Networks,
Vol.\ 6, No.\ 12,
Dec. 2013,
pp.\ 1590-1605,
\doi{10.1002/sec.166}.

\bibitem{iocheck}
F.Zhang,
``IOCheck: A framework to enhance the security of I/O devices at runtime'',
43rd Annual IEEE/IFIP Conference on Dependable Systems and Networks Workshop (DSN-W),
Budapest (Hungary), June 24-27, 2013,
\doi{10.1109/DSNW.2013.6615523}.

\bibitem{raspberry}
Raspberry Pi Foundation,
``Raspberry Pi'',
\url{https://www.raspberrypi.org/}.

\bibitem{symbiote_web}
Red ballon security,
``Symbiote Defense'',
\url{http://www.redballoonsecurity.com}.

\bibitem{autoscopy_web}
Trustworthy Cyber Infrastructure for the Power Grid,
``Autoscopy Jr.'',
\url{https://tcipg.org/technology/autoscopy-jr}.

\bibitem{bcm2835}
Raspberry Pi Foundation,
``BCM2835'',
\url{https://www.raspberrypi.org/documentation/hardware/raspberrypi/bcm2835/README.md}.

\bibitem{codesys_runtime}
3S-Smart Software Solutions GmbH,
``CODESYS Control for Raspberry Pi SL'',
\url{http://store.codesys.com/codesys-control-for-raspberry-pi-sl.html}.

\bibitem{codesys_dev}
3S-Smart Software Solutions GmbH,
``CODESYS Development System V3'',
\url{http://store.codesys.com/codesys.html}.

\bibitem{wago_linux}
WAGO Kontakttechnik GmbH \& Co.,
``Linux - Automation for the Future'',
\url{http://global.wago.com/en/products/new-items/overview/basic-page-2600.jsp}.

\bibitem{am35x}
Texas Instruments,
``AM3517 Sitara Processor Technical documents'',
\url{http://www.ti.com/product/AM3517/technicaldocuments}.

\bibitem{strace}
``strace: Linux syscall tracer'',
\url{https://strace.io/}.

\bibitem{kprobes}
J.Keniston, P.S.Panchamukhi, M.Hiramatsu,
``Kernel probes'',
\url{https://www.kernel.org/doc/Documentation/kprobes.txt}.

\end{thebibliography}
