\documentclass[pdfa,cucitura]{toptesi}

\hypersetup{
    pdfpagemode={UseOutlines},
    bookmarksopen,
    pdfstartview={FitH},
    colorlinks,
    linkcolor={blue},
    citecolor={red},
    urlcolor={blue}
}

%\linespread{1.3}

\input{commands.tex}

\begin{document}
\selectlanguage{english}

\CorsoDiLaureaIn{Corso di Laurea Magistrale in }
\TesiDiLaurea{}
\AdvisorName{Supervisor}
\CandidateName{Candidate}


\logosede{res/logopolito}
\ateneo{Politecnico di Torino}

\titolo{Analysis and detection of I/O attacks on Programmable Logic Controllers}

\corsodilaurea{Ingegneria Informatica}

\candidato{Andrea \textsc{Genuise}}

\relatore{prof.\ Antonio Lioy}

\sedutadilaurea{\textsc{Academic~Year} 2016-2017}


\errorcontextlines=9

\frontespizio
\paginavuota
\newpage

\advance\voffset -5mm
\advance\textheight 30mm


\begin{dedica}
Alla mia famiglia

\textdagger\ A nonno Carmelo
\end{dedica}


\sommario

Recently embedded systems have become more and more integrated with several aspects of our lives, and their security concerns have risen as well.
In particular, Programmable Logic Controllers (PLCs), which are embedded systems deployed within the context of an Industrial Control System (ICS),
need to be reliable and secure because they can have direct effects on critical processes.
From an engineering perspective, they are connected to sensors and actuators and use Input/Output (I/O) interfaces to interact with the physical world.
As demonstrated by a novel kind of attack, called Pin Control Attack or I/O attack, a malicious user can tamper with the integrity or
the availability of legitimate I/O operations, possibly causing physical damage to people and environment.
In this thesis, we first analyse this new threat and discuss its applicability on real PLCs, then we design and implement a detection system able to raise the bar for the attacker.
The implementation, provided as a Linux kernel module, is evaluated on an ARM-based architecture,
showing its effectiveness and impact on PLCs, which typically have very limited resources and strict timing requirements.


\ringraziamenti

This thesis is the result of the work conducted as guest student at University of Twente, in the Netherlands. I want to thank my supervisor, prof. Antonio Lioy,
to have authorised this collaboration and allowed me to have this experience abroad, which has been a fundamental educational experience both from professional and human perspectives.

I am particularly grateful to prof. Sandro Etalle for his care in supervising my work,
and to Ali Abbasi for his continuous support and advices during all phases of my project.
I would like to thank the Services, Cybersecurity and Safety research group of University of Twente for accepting me as guest student
and allowing me to be part of it for 6 months.

Thanks to all friends I met in the Netherlands who have made it an amazing experience, for making me feel like part of a big international family.
Thanks also to all my friends known during my master's degree, with whom I shared these 2 years of my life.

I wish to thank my family, who always believes in me and supports my life choices, even if it means living far away from me.

Finally, my special thanks are for Giulia, who has shared this whole experience with me from the beginning, both in good and bad moments.
She has always been my strength, and without her it would not have been the same adventure.

\hfill My sincere thanks,

\hfill Andrea

\selectlanguage{italian}

\ringraziamenti

Questa tesi \`e il risultato del lavoro svolto come studente ospite di University of Twente, nei Paesi Bassi. Voglio ringraziare il mio relatore, prof. Antonio Lioy,
per aver autorizzato questa collaborazione e avermi permesso di svolgere la tesi all'estero, che si \`e dimostrata un esperienza formativa fondamentale,
sia dal punto di vista professionale che umano.

Sono particolarmente grato al prof. Sandro Etalle per la sua cura nella supervisione del mio lavoro,
e ad Ali Abbasi per il suo continuo supporto e i suoi consigli durante tutte le fasi del mio progetto.
Voglio ringraziare il gruppo di ricerca Services, Cybersecurity and Safety di University of Twente per avermi accettato come studente ospite
e avermi permesso di essere parte di esso per 6 mesi.

Grazie a tutti gli amici incontrati nei Paesi Bassi che hanno reso tutto ciò un'esperienza indimenticabile, per avermi fatto sentire parte di una grande famiglia internazionale.
Grazie a tutti i miei amici conosciuti durante l'intero percorso di laurea magistrale, con cui ho condiviso questi 2 anni della mia vita.

Voglio ringraziare la mia famiglia, che crede sempre in me e sostiene ogni mia scelta di vita, anche se comporta continui sacrifici come vivere a distanza.

Infine, un ringraziamento speciale a Giulia, che ha condiviso con me quest'esperienza dall'inizio alla fine, sia nei momenti di gioia che di tristezza.
Lei \`e sempre stata la mia forza, e senza di lei non sarebbe stata la stessa avventura.

\hfill Grazie di cuore,

\hfill Andrea

\selectlanguage{english}

%\tablespagetrue
%\figurespagetrue

\indici

\mainmatter


\input{introduction}

\input{background}

\input{attack}

\input{defense}

\input{results}

\input{conclusions}


\input{biblio.tex}


%\begin{appendices}
%\setcounter{table}{0}
%\setcounter{figure}{0}
%\renewcommand\thetable{\thesection\arabic{table}}
%\renewcommand\thefigure{\thesection\arabic{figure}}

%\end{appendices}


\end{document}
