\begin{thebibliography}{99}

\bibitem{ghostplc}
A.Abbasi, M.Hashemi, E.Zambon, S.Etalle,
``Stealth Low-Level Manipulation of Programmable Logic Controllers I/O By Pin Control Exploitation'',
TODO CRITIS Conference 2016.

\bibitem{stuxnet}
N.Falliere, L.O Murchu, E.Chien,
``W32.Stuxnet  Dossier'',
Version 1.4, Febr. 2011.
Online: \url{http://www.symantec.com/content/en/us/enterprise/media/security_response/whitepapers/w32_stuxnet_dossier.pdf}.

\bibitem{io-command}
R.Grandgenett, W.Mahoney, R.Gandhi,
``Authentication Bypass and Remote Escalated I/O Command Attacks'',
CISR '15: Proceedings of the 10th Annual Cyber and Information Security Research Conference,
Oak Ridge, Tennesse (USA), April 7-9, 2015,
\doi{10.1145/2746266.2746268}.

\bibitem{taxonomy}
D.Papp, Z.Ma, L.Buttyan,
``Embedded systems security: Threats, vulnerabilities, and attack taxonomy''
Privacy, Security and Trust (PST) 13th Annual Conference,
Izmir (Turkey), July 21-23, 2015
pp.\ 145-152,
\doi{10.1109/PST.2015.7232966}.

\bibitem{firmware-mod}
Z.Basnight, J.Butts, J.Lopez Jr., T.Dube,
``Firmware modification attacks on programmable logic controllers'',
International Journal of Critical Infrastructure Protection,
Volume 6, Issue 2, June 2013,
pp.\ 76-84,
\doi{10.1016/j.ijcip.2013.04.004}.

\bibitem{ethernet-vuln}
D.Peck, D.Peterson,
``Leveraging ethernet card vulnerabilities in field devices'',
Proceedings of the SCADA Security Scientific Symposium,
Miami Beach, Florida (USA), Jan. 18-19, 2009,
pp.\ 1-19.

\bibitem{print-vuln}
A.Cui, M.Costello, S.J.Stolfo,
``When Firmware Modifications Attack: A Case Study of Embedded Exploitation'',
20th Annual Network \& Distributed System Security Symposium,
San Diego, California (USA), Febr. 24-27, 2013,
\doi{10.7916/D8P55NKB}.

\bibitem{dynamic-payload}
S.McLaughlin,
``On Dynamic Malware Payloads Aimed at Programmable Logic Controllers'',
In 6th USENIX Workshop on Hot Topics in Security,
2011.

\bibitem{sabot}
S.McLaughlin, P.McDaniel,
``SABOT: Specification-based Payload Generation for Programmable Logic Controllers'',
CCS '12: Proceedings of the 2012 ACM conference on Computer and Communications Security,
New York (USA), Oct. 16-18, 2012,
pp.\ 439-449,
\doi{10.1145/2382196.2382244}.

\bibitem{siemens-s7}
D.Beresford,
``Exploiting Siemens Simatic S7 PLCs'',
Black Hat USA 2011,
Las Vegas, Nevada (USA), Aug. 3-4, 2011.

\bibitem{plc-network}
J.Klick, S.Lau, D.Marzin, J.Malchow, V.Roth,
``Internet-facing PLCs as a Network Backdoor'',
Proceedings of IEEE Conference on Communications and Network Security (CNS),
Florence (Italy), Sept. 28-30, 2015,
pp.\ 524-532,
\doi{10.1109/CNS.2015.7346865}.

\bibitem{schneider-bof}
ICS-CERT,
``Schneider Electric Modicon M340 Buffer Overflow Vulnerability'',
Dec. 17, 2015.
Online: \url{https://ics-cert.us-cert.gov/advisories/ICSA-15-351-01}.

\bibitem{rockwell-vuln}
ICS-CERT,
``Rockwell Automation Micrologix 1100 and 1400 PLC Systems Vulnerabilities (Update A)'',
Oct. 27, 2015.
Online: \url{https://ics-cert.us-cert.gov/advisories/ICSA-15-300-03A}.

\bibitem{rockwell-vuln2}
ICS-CERT,
``Rockwell Automation MicroLogix 1100 PLC Overflow Vulnerability'',
Jan. 26, 2016.
Online: \url{https://ics-cert.us-cert.gov/advisories/ICSA-16-026-02}.

\bibitem{elcsoft-vuln}
ICS-CERT,
``Eaton ELCSoft Programming Software Memory Vulnerabilities'',
June 30, 2016.
Online: \url{https://ics-cert.us-cert.gov/advisories/ICSA-16-182-01}.

\bibitem{jop}
P.Chen, X.Xing, B.Mao, L.Xie,
``Return-oriented rootkit without returns (on the x86)''
ICICS 2010: 12th International Conference on Information and Communications Security,
Barcelona (Spain), Dec. 15-17, 2010,
pp.\ 340-354,
\doi{10.1007/978-3-642-17650-0_24}.

\bibitem{no-ret}
S.Checkoway, L.Davi, A.Dmitrienko, A.Sadeghi, H.Shacham, M.Winandy,
``Return-Oriented Programming without Returns'',
CCS '10: Proceedings of the 17th ACM conference on Computer and Communications Security,
Chicago, Illinois (USA), Oct. 4-8, 2010,
\doi{10.1145/1866307.1866370}.


\end{thebibliography}
